---
title: "Project InfoVoz"
author: "Dorina Craciun, Lazar Alexandru"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# World Happinesss Report

Greetings! In this project we decided to explore World Happiness dataset in order to find what are the key factors that keep people happy and communities together. We think it's a great idea to see the differences between regions in **why** are some regions happier than others and **what** is to be done in order to balance things given the fact that we live in times when we should all embrace each other and keep our heads up.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
