---
title: "Project InfoVoz"
author: "Dorina Craciun, Lazar Alexandru"
output: html_document
---
<style type="text/css">
  body{
  font-size: 16pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# World Happinesss Report

Greetings! In this project we decided to explore World Happiness dataset in order to find what are the key factors that keep people happy and communities together. We think it's a great idea to see the differences between regions: in **why** are some regions happier than others and **what** is to be done in order to balance things given the fact that we live in times when we should all embrace each other and keep our heads up.

### Exploration of the data set

For the first part let's explore a bit the data set and explore some interesting facts.  

```{r include=FALSE}
library(tidyverse) 
library(reshape2)
library(lattice)
library(psych)
library(DataExplorer)
library(viridis)

#reading data
df15 <- read_csv("input/2015.csv")
df16 <- read_csv("input/2016.csv")
df17 <- read_csv("input/2017.csv")
df18 <- read_csv("input/2018.csv")
df19 <- read_csv("input/2019.csv")

df15 <- df15 %>% mutate(Year = 2015)
df16 <- df16 %>% mutate(Year = 2016)
df17 <- df17 %>% mutate(Year = 2017)
df18 <- df18 %>% mutate(Year = 2018)
df19 <- df19 %>% mutate(Year = 2019)

# rename the columns to standardize
df17 <- df17 %>% rename(c("Happiness Rank" = "Happiness.Rank" ,              
                          "Happiness Score" = "Happiness.Score",
                          "Economy (GDP per Capita)" = "Economy..GDP.per.Capita.",  
                          "Health (Life Expectancy)" = "Health..Life.Expectancy.",     
                          "Trust (Government Corruption)" = "Trust..Government.Corruption.",
                          "Dystopia Residual" = "Dystopia.Residual"))

df18 <- df18 %>% rename(c("Country" = "Country or region",
                          "Happiness Rank" = "Overall rank",
                          "Happiness Score" = "Score",
                          "Economy (GDP per Capita)" = "GDP per capita",              
                          "Health (Life Expectancy)" = "Healthy life expectancy",
                          "Freedom" = "Freedom to make life choices" ,
                          "Trust (Government Corruption)" = "Perceptions of corruption")) 

# chart to num
df18$`Trust (Government Corruption)` <-  as.numeric(df18$`Trust (Government Corruption)`)

df19 <- df19 %>% rename(c("Country" = "Country or region",
                          "Happiness Rank" = "Overall rank",
                          "Happiness Score" = "Score",
                          "Economy (GDP per Capita)" = "GDP per capita",
                          "Health (Life Expectancy)" = "Healthy life expectancy",    
                          "Freedom" = "Freedom to make life choices",
                          "Trust (Government Corruption)" = "Perceptions of corruption"))
# bind tables
a <- bind_rows(df15, df16)
b <- bind_rows(a, df17)
c <- bind_rows(b, df18)
d <- bind_rows(c, df19)

n <- c("Family", "Standard Error", "Dystopia Residual", "Region",  
       "Upper Confidence Interval", "Lower Confidence Interval",
       "Whisker.high", "Whisker.low", "Social support")
d <- d[ , !(names(d) %in% n)]


r1 <- df15 %>% select("Country", "Region")
r2 <- df16 %>% select("Country", "Region")
r3 <- full_join(r1, r2, by = c("Country" = "Country","Region" = "Region"))

df <- left_join(d, r3, by = "Country"); head(df); tail(df)

df <- df %>% drop_na()
df$Region <- as.factor(df$Region)
df$Year <- as.factor(df$Year)
df <- df %>% mutate(Region = fct_recode(Region,"Southeastern Asia" = "Southern Asia", "Southeastern Asia" = "Eastern Asia"))
```

Let's check how many countries each region have. It's an important factor in calculating the mean of happiness for those regions.  

```{r echo=FALSE}
ggplot(df, aes(Region, fill = Region)) +
  geom_bar(stat = "count", show.legend = F, color = 'black') + 
  coord_flip() +
  labs(x = "", fill = "Region", y = "No. of countries", 
       title = "No. of countries per region") + 
  theme_light(base_size = 18)

```

Let's calculate now the meam and the median of happiness score for each region.
```{r , echo=FALSE, warning=FALSE, message=FALSE}
# Compute descriptive statistics by groups

library(ggpubr)
stable <- desc_statby(df, measure.var = "Happiness Score",
                      grps = "Region")
stable <- stable[, c("Region","mean","median")]
names(stable) <- c("Region", "Mean of happiness score","Median of happiness score")
# Summary table plot
stable.p <- ggtexttable(stable,rows = NULL, 
                        theme = ttheme("classic"))

stable.p  
```

Now let's check the mean of some important features over the regions.

```{r figures-side, fig.show="hold", out.width="50%", echo=FALSE}
ss <- df %>%
     select(Region, `Happiness Score`) %>%
     group_by(Region) %>%
     summarise(n = n(), mn = mean(`Happiness Score`), md = median(`Happiness Score`), std = sd(`Happiness Score`))

ggplot(ss, aes(x = fct_reorder(Region, md), y = mn, fill = mn)) +
  geom_bar(stat = "Identity", show.legend = F) +
  scale_fill_gradient(low = "red", high = "green") +
  labs(x = "", y = "Median happiness score", fill = "Score",
       title = "Median happiness score") +
  theme_light(base_size = 18) + coord_flip()

ss <- df %>%
     select(Region, `Economy (GDP per Capita)`) %>%
     group_by(Region) %>%
     summarise(n = n(), md = median(`Economy (GDP per Capita)`))

ggplot(data = ss, aes(x = Region, y = md, fill = md)) + 
  geom_bar(stat = "identity", show.legend = F) + 
  scale_fill_gradient(low = "red", high = "green") +
  labs(x= "", y = "Economy score", 
       title = "Economy") + 
  theme_light(base_size = 18) + coord_flip()
```

```{r fig.show="hold", out.width="50%", echo=FALSE}
ss <- df %>% select(Region, `Health (Life Expectancy)`) %>%
  group_by(Region) %>%
  summarise(md = median(`Health (Life Expectancy)`))

ggplot(data = ss, aes(x = Region, y = md, fill = md)) +
  geom_bar(stat = "identity", show.legend = F) +
  scale_fill_gradient(low = "red", high = "green") +
  labs(x = "", y = "Life expectancy percentage", 
       title = "Life expectancy") + 
  coord_flip() + theme_light(base_size = 18)


ss <- df %>% select(Region, Freedom) %>%
  group_by(Region) %>%
  summarise(md = median(Freedom))

ggplot(data = ss, aes(x = Region, y = md, fill = md)) +
  geom_bar(stat = "identity", show.legend = F) + 
  scale_fill_gradient(low = "red", high = "green") + 
  labs(x = "", title = "Freedom", y = "Freedom percentage") +
  coord_flip() + theme_light(base_size = 18)
```

```{r fig.show="hold", out.width="50%", echo=FALSE}
ss <- df %>%
      select(Region, `Trust (Government Corruption)`) %>%
      group_by(Region) %>%
      summarise(md = median(`Trust (Government Corruption)`))

ggplot(data = ss, aes(x = Region, y = md, fill = md)) +
  geom_bar(stat = "identity", show.legend = F) +
  scale_fill_gradient(low = "red", high = "green") +
  labs(x = "", title = "Trust in governmnent", 
       y = "Trust score") +
  coord_flip() + theme_light(base_size = 18)
```

### Happiness Score Around the World
```{r, echo=FALSE, warning=FALSE, message=FALSE}

w <- map_data("world")
colnames(w)[5]<- "Country"
myw <- inner_join(w, df, by = "Country")
worldplot <- ggplot(data = w, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray") +
  geom_polygon(data=myw, aes(x = long, y = lat, group = group, fill = `Happiness Score`),color = "white") +
  geom_polygon(color = "black", fill = NA) +
  theme_bw() +
  ggtitle("Happiness Score in the World") +
  scale_fill_distiller(palette = "Spectral")
worldplot

```



### Interactive plot Happiness vs GDP
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# install.packages("plotly")
library(plotly)
plot_ly(df ,x = ~`Economy (GDP per Capita)`, y = ~`Happiness Score`,  hoverinfo = "text", text = ~Country) %>%
  add_markers(frame = ~Region, size = ~`Happiness Score`, color = ~Region,
              marker = list(sizemode = "diameter",  sizeref = 3)) %>%
  animation_opts(
    frame = 2000, 
    transition = 300, 
    easing = "elastic"
  ) %>%
  animation_slider(
    currentvalue = list(
      prefix = NULL, 
      font = list(color = "red")
    )
  ) %>%   
  layout(
    xaxis = list(title = "GDP per Capita"),
    yaxis = list(title = "Happiness Score"),
    title = "Happiness vs GDP")
```



### Interactive plot 2 Happiness vs Economy
```{r, echo=FALSE, warning=FALSE, message=FALSE}


colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')
plot_ly(df, x = ~`Economy (GDP per Capita)`, y = ~`Happiness Score`, z = ~`Health (Life Expectancy)`, color = ~Region, size = ~`Happiness Score`, colors = colors,
        marker = list(symbol = 'circle', sizemode = 'diameter'), sizes = c(2, 20),
        text = ~paste('Country:', Country, '<br>Life Expectancy:', `Health (Life Expectancy)`, '<br>Happiness Score:', `Happiness Score`,
                      '<br>GDP.:', `Economy (GDP per Capita)`)) %>%
  layout(title = 'Happiness v. Per Capita GDP',
         scene = list(xaxis = list(title = 'GDP)',
                                   gridcolor = 'rgb(255, 255, 255)',
                                   zerolinewidth = 1,
                                   ticklen = 5,
                                   gridwidth = 2),
                      yaxis = list(title = 'Happiness Score',
                                   gridcolor = 'rgb(255, 255, 255)',
                                   zerolinewidth = 1,
                                   ticklen = 5,
                                   gridwith = 2),
                      zaxis = list(title = 'Life expectancy',
                                   gridcolor = 'rgb(255, 255, 255)',
                                   zerolinewidth = 1,
                                   ticklen = 5,
                                   gridwith = 2)),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')


```



### Correlation plot
Let’s see the correlation between numerical variables in our dataset.
According to this cor plot, Economy, life expectancy, and freedom play the most significant role in contributing to happiness. Trust and generosity have the lowest impact on the happiness score.
```{r , echo=FALSE, warning=FALSE, message=FALSE}
library(ggcorrplot)
# Compute a correlation matrix
cor_df <- subset(df, select = -c(1, 2, 9, 10))

corr <- round(cor(cor_df), 1)

# Visualize the correlation matrix
ggcorrplot(corr, 
           # hc.order = TRUE, #for ordering; using hierarchical clustering
           #type = "lower", #for lower triangle
           lab = TRUE,
           lab_size = 4,
           title = "Correlation of Variables",
           ggtheme = theme_bw)
```

### Comparing different continents regarding their happiness variables
Let’s calculate the average happiness score and the average of the other variables for each continent. Then melt it to have variables and values in separate columns. Finally, using ggplot to show the difference between continents.

```{r , echo=FALSE, warning=FALSE, message=FALSE}
df.Region <- df %>%
  select(-3) %>%
  group_by(Region) %>%
  summarise_at(vars(-Country, -`Happiness Rank`, -Year), funs(mean(.)))

df.Region.melt <- melt(df.Region)

ggplot(df.Region.melt, aes(y=value, x=Region, color=Region, fill=Region)) + 
  geom_bar( stat="identity") +    
  facet_wrap(~variable) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Average value of happiness variables for different regions", 
       y = "Average value") 

```

We can see that Australia has approximately the highest average in all fields except economy, after that North America and Western Europe are roughly the same regarding happiness score and the other factors. Finally, Sub Saharan Africa has the lowest scores in all fields.



### Scatter plot with regression line
Let’s see the correlation between happiness score and life expectancy in the happiness dataset for different continents by creating a scatter plot.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df, aes(x = `Health (Life Expectancy)`, y = `Happiness Score`)) + 
  geom_point(aes(color=Region), size = 3, alpha = 0.8) +  
  geom_smooth(aes(color = Region, fill = Region), 
              method = "lm", fullrange = TRUE) +
  facet_wrap(~Region) +
  theme_bw() + labs(title = "Scatter plot with regression line")
```


The correlation between life expectancy and happiness score in Europe, North America, and Asia is more significant than the other continents. 



### Happiness score comparison on different regions 
Next let's see a comparison of the happiness score on different regions by creating 2 types of plots.

#### Box plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df , aes(x = Region, y = `Happiness Score`)) +
  geom_boxplot(aes(fill=Region)) + theme_bw() +
  theme(axis.title = element_text(size = (8)), axis.text.x = element_text(angle = 90, hjust = 1))
```

#### Violin plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df, aes(x=Region, y=`Happiness Score`))+
  geom_violin(aes(fill=Region),alpha=0.7,scale = "width")+ theme_bw() +
  theme(axis.title = element_text(size = (8)), axis.text.x = element_text(angle = 90, hjust = 1))
```

### Linear Regression for predicting the happines of a country judging by the health/life expectancy
For the first model we tried **Linear Regression** to try to predict the happiness based on the what looks to be the best predictor: Life expectancy

```{r , echo=FALSE, warning=FALSE, message=FALSE}

Linear_Model_1 = lm(df$`Happiness Score`~df$`Health (Life Expectancy)`, data = df)
Linear_Model_1
#hapiness_score = 3.303+3.382*health


# The correlation between the Health Expectancy and the Happiness score
cor(df$`Happiness Score`, df$`Health (Life Expectancy)`)

summary(Linear_Model_1)
summary(Linear_Model_1)$coefficients
standard_residuals_1 = rstandard(Linear_Model_1) #Standardized residuals"
resid_1 = Linear_Model_1$residuals
# predict_1 = predict(Linear_Model_1)

# standard_residuals_1
# resid_1


ggplot(df, aes(x=`Health (Life Expectancy)`,y=`Happiness Score`)) + 
  geom_point() + 
  geom_smooth(method="lm") +
  labs(x="Health Expectancy",y="Happiness Score") + 
  geom_hline(yintercept=mean(df$`Happiness Score`),size=0.5) +
  geom_vline(xintercept=mean(df$`Health (Life Expectancy)`),size=0.5) +
  ggtitle("Linear Regression predicting hapiness score using the health/life expentacy")


# Residuals
ggplot(data = Linear_Model_1, aes(x=.fitted, y = .resid)) + 
  geom_point()+
  labs(x="Health Expectancy Predictions",y="Residuals") 
  
library(ggfortify)
autoplot(Linear_Model_1)
```

### Multiple linear regression model with all parameters included.

```{r , echo=FALSE, warning=FALSE, message=FALSE}
# Splitting the dataset into the Training set and Test set
library(caTools)
dataset <- df[2:10]
split = sample.split(dataset$`Happiness Score`, SplitRatio = 0.8)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)

Linear_Model_4 <- lm(formula = `Happiness Score` ~ .,
                     data = training_set)
summary(Linear_Model_4)

predict_4 = predict(Linear_Model_4, newdata = test_set)

predict_actual_4 <- as.data.frame(cbind(Prediction = predict_4, Actual = test_set$`Happiness Score`))

ggplot(predict_actual_4, aes(Actual, Prediction )) +
  geom_point() + geom_abline() +
  labs(title = "Multiple Linear Regression", x = "Actual happiness score",
       y = "Predicted happiness score") 
```


### Decisional Tree


```{r}
dataset <- df[-1]
dataset$`Happiness Rank` <- NULL
dataset$Year <- NULL
dataset$Region <- NULL
split = sample.split(dataset$`Happiness Score`, SplitRatio = 0.8)
training_dt = subset(dataset, split == TRUE)
test_dt = subset(dataset, split == FALSE)
library(rpart)
dt = rpart(formula = training_dt$`Happiness Score` ~ .,
                  data = training_dt,
                  control = rpart.control(minsplit = 10))

y_pred_dt = predict(dt, newdata = test_dt)

actual_pred_dt <- as.data.frame(cbind(Prediction = y_pred_dt, Actual = test_dt$`Happiness Score`))


gg.dt <- ggplot(actual_pred_dt, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Decisional Tree", x = "Actual happiness score",
       y = "Predicted happiness score")
gg.dt


# Plotting the tree
library(rpart.plot)
prp(dt)
```

### Support Vector Regression

```{r include=FALSE}
library(caTools)
set.seed(123)
data_svr <- df[-1]
split = sample.split(data_svr$`Happiness Score`, SplitRatio = 0.8)
training = subset(data_svr, split == TRUE)
test = subset(data_svr, split == FALSE)
```


```{r}
library(e1071)
svr = svm(formula = training$`Happiness Score` ~ .,
                data = training,
                type = 'eps-regression',
                kernel = 'radial')

y_pred_svr = predict(svr,  newdata = test)

Pred_Actual_svr <- as.data.frame(cbind(Prediction = y_pred_svr, Actual = test$`Happiness Score`))




gg.svr <- ggplot(Pred_Actual_svr, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "SVR", x = "Actual happiness score",
       y = "Predicted happiness score")

gg.svr

```

### Random Forest Regression

```{r}
library(randomForest)
#split up sample

dataset <- df[-1]
split = sample.split(dataset$`Happiness Score`, SplitRatio = 0.7)

#70% training data, 30% test data

train_rf <- subset(dataset, split==TRUE)
test_rf <- subset(dataset, split==FALSE)

rf.model <- randomForest(x=train_rf, y=train_rf$`Happiness Score`, importance = TRUE)
rf.model

importance(rf.model, type = 2)

print('The higher value indicates that the variable is more important than others. We can tell that Economy has the highest number, which is the most significant, and Generosity is the least important.')

#INTERPRET THE MODEL
print(summary(rf.model))

res_rf <- residuals(rf.model)
class(res_rf)


#PREDICTIONS

Score.predictions_rf <- predict(rf.model, test_rf)

results_rf <-cbind(Score.predictions_rf,test_rf$`Happiness Score`)
colnames(results_rf) <-c('predicted','actual')
results_rf<-as.data.frame(results_rf)

print(head(results_rf))

#MSE

mse_rf <- mean( (results_rf$actual - results_rf$predicted)^2)
print("MSE RANDOM FOREST")
print(mse_rf)

#RMSE
print("RMSE RANDOM FOREST")
print(mse_rf^0.5)


```

```{r , echo=FALSE, warning=FALSE, message=FALSE}

# install.packages("randomForest")
library(randomForest)

rf <- randomForest(x = dataset[-1],
                            y = dataset$`Happiness Score`,
                            ntree = 500)

predict_rf = predict(rf, newdata = test_set)

predict_actual_rf <- as.data.frame(cbind(Prediction = predict_rf, Actual = test_set$`Happiness Score`))


ggplot(predict_actual_rf, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Random Forest Regression", x = "Actual happiness score",
       y = "Predicted happiness score")
```


### Neuralnet
```{r , echo=FALSE, warning=FALSE, message=FALSE}
plot(pressure)
```
<!-- Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. -->
